---
title: "ES Toolkit Examples: Heavy Precipitation (CSP1) and Growing Degree Days (CST3) Metrics"
author: "Lisa Nelson - IMD"
date: "February 12, 2018"
output: html_document
highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EnvironmentalSettingToolkit)
library(ggplot2)
```

## IMD Environmental Setting Toolkit

The EnvironmentalSetting toolkit <https://github.com/nationalparkservice/EnvironmentalSetting_Toolkit> accesses climate stations and observation data for areas around and within National Parks. These data are vended by webservices maintained by NOAA's Regional Climate Centers (RCCs) from a system called ACIS - Applied Climate Information System <http://www.rcc-acis.org/index.html>. This example shows two of the 52 climate metrics in the draft Environmental Setting protocol uses.

Select a park, then decide which climate parameters to search.  The findStation function returns a data frame of station 'metadata', in this case for stations recording the growing degree day parameter within 10km of Shenandoah National Park:

```{r findStation, message=FALSE, warning=FALSE, error=FALSE}
head(findStation(unitCode = "SHEN", distance=10, climateParameters=list("gdd")))
```

## Getting weather observations

The set of stations can be used to request weather observations for a climate parameter, specifying only those stations that are U.S. Historical Climatology Network stations that collected data within the past week:

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
# Get stations collecting precipitation data
stationsSHENCSP1 <- (findStation(unitCode = "SHEN", distance=10, climateParameters=list("pcpn")))
# Subset HCN stations that collected data up to the first full week of March 2018 
hcnStationsSHENCSP1 <- stationsSHENCSP1[stationsSHENCSP1$isHCNStation == "Y" & stationsSHENCSP1$maxDate >= "2018-03-10",]
hcnStationsSHENCSP1
```
This subset of stations (hcnStationsSHENCSP1) is used to request the annual count of precipitation events greater than or equal to 1.0 inch.
```{r message=FALSE, warning=FALSE, error=FALSE, echo=TRUE}
# Get observations exceeding 1.0 inch of precipitation since 2000 - only head rows shown for brevity. 
head(getWxObservations(climateStations = hcnStationsSHENCSP1, climateParameters = list("pcpn"), duration="yly", interval="yly", reduceCodes = list("cnt_ge_1.0"), sdate = "2000-01-01"))
```

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
# Get stations collecting growing degree day data
stationsSHENGDD <- (findStation(unitCode = "SHEN", distance=10, climateParameters=list("gdd")))
# Subset HCN stations that collected data up to the first full week of March 2018 
hcnStationsSHENGDD <- stationsSHENGDD[stationsSHENGDD$isHCNStation == "Y" & stationsSHENGDD$maxDate >= "2018-03-10",]
hcnStationsSHENGDD
```
This subset of stations (hcnStationsSHENGDD) is used to request the annual count (sum) of growing degree days greater than 0.
```{r message=FALSE, warning=FALSE, error=FALSE, echo=TRUE}
# The protocol uses a base of 32 degrees for growing degree data. Get observations since 2000 - only head rows shown for brevity. 
head(getWxObservations(climateStations = hcnStationsSHENGDD, climateParameters = list("gdd32"), duration="yly", interval="yly", reduceCodes = list("cnt_gt_0"), sdate = "2000-01-01"))
```
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
obsDataCSP1 <- getWxObservations(climateStations = hcnStationsSHENCSP1, climateParameters = list("pcpn"), duration="yly", interval="yly", reduceCodes = list("cnt_ge_1.0"), sdate = "2000-01-01")

obsDataGDD <- getWxObservations(climateStations = hcnStationsSHENGDD, climateParameters = list("gdd32"), duration="yly", interval="yly", reduceCodes = list("cnt_gt_0"), sdate = "2000-01-01")
```
Note the Toolkit does not include graphs - these are included in this notebook to offer a quick visual check.  The last two stations have a longer time series for the gdd parameter for the specified date range. 

```{r echo=FALSE, dpi=96, fig.width=8, fig.height=7, message=FALSE, warning=FALSE, error=FALSE}
ggplot(data = obsDataCSP1, aes(x=obsDataCSP1$date, y=obsDataCSP1$pcpn_in_cnt_ge_1.0, group=obsDataCSP1$name, color=obsDataCSP1$name)) + geom_line() + geom_point() + labs(x = "Year", y = "# days with pcpn >= 1.0 inch", title = "CSP1: Heavy Precipitation", color = "Station")

ggplot(data = obsDataGDD[1:75], aes(x=obsDataGDD$date[1:75], y=obsDataGDD$gdd32__cnt_gt_0[1:75], group=obsDataGDD$name[1:75], color=obsDataGDD$name[1:75])) + geom_line() + geom_point() + labs(x = "Year", y = "# growing degree days (annual sum)", title = "CST6: Growing Degree Days (annual sum)", color = "Station")

```
