---
title: "ES Toolkit Examples"
author: "Lisa Nelson - IMD"
date: "February 12, 2018"
output: html_document
highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EnvironmentalSettingToolkit)
```

## IMD Environmental Setting Toolkit

The EnvironmentalSetting toolkit <https://github.com/nationalparkservice/EnvironmentalSetting_Toolkit> accesses climate stations and observation data for areas around and within National Parks. These data are vended by webservices maintained by NOAA's Regional Climate Centers (RCCs) from a system called ACIS - Applied Climate Information System <http://www.rcc-acis.org/index.html>.

Select a park, then decide which climate parameters to search.  The findStation function returns a data frame of station 'metadata', in this case for stations recording the growing degree day parameter within 10km of Shenandoah National Park:

```{r findStation, message=FALSE, warning=FALSE, error=FALSE}
head(findStation(unitCode = "SHEN", distance=10, climateParameters=list("gdd")))
```

## Getting weather observations

The set of stations can be used to request weather observations for a climate parameter, specifying only those stations that are U.S. Historical Climatology Network stations that collected data within the past week:

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
# Get stations collecting growing degree day
stationsSHEN <- (findStation(unitCode = "SHEN", distance=10, climateParameters=list("gdd")))
# Subset HCN stations that collected data up to the first week of February 2018 
hcnStationsSHEN <- stationsSHEN[stationsSHEN$isHCNStation == "Y" & stationsSHEN$maxDate >= "2018-02-10",]
hcnStationsSHEN
```
This subset of stations (hcnStationsSHEN) is used to request the annual sum of growing degree days greater than 0.
```{r message=FALSE, warning=FALSE, error=FALSE, echo=TRUE}
# Get observations since 2000 - only head rows for brevity 
head(getWxObservations(climateStations = hcnStationsSHEN, climateParameters = list("gdd"), duration="yly", reduceCodes = list("cnt_gt_0"), sdate = "2000-01-01"))
```
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
obsData <- NULL
obsData <- getWxObservations(climateStations = hcnStationsSHEN, climateParameters = list("gdd"), duration="yly", reduceCodes = list("cnt_gt_0"), sdate = "2000-01-01")
```
Note the Toolkit does not include graphs - these are included in this notebook to offer a quick visual check.  The last two stations have a longer time series for the gdd parameter for the specified date range. 

```{r echo=FALSE, dpi=96, fig.width=8, fig.height=7, message=FALSE, warning=FALSE, error=FALSE}
# Iterate observations, graphing a scatter plot for each station
plot.new()
par(mfrow=c(length(unique(obsData$uid)),1)) # All station graphs in one page column
par(mar=c(4.5,4.5,1.5,1.5)) # Graph margins in lines
for (s in unique(obsData$uid)) {
  #par(mar=c(4.5,4.5,1.5,1.5))
  # x = obsData$date[which(obsData$uid == s)]
  # y = obsData$gdd__cnt_gt_0[which(obsData$uid == s)]
  # m = obsData$name[which(obsData$uid == s)][1]
  plot(obsData$date[which(obsData$uid == s)], obsData$gdd__cnt_gt_0[which(obsData$uid == s)], pch=16, col="blue", ylab="count (annual sum)", xlab="date", main=obsData$name[which(obsData$uid == s)][1])
}
```
