---
title: "Environmental Setting Metrics - Example ACIS Calls"
author: "Lisa Nelson - IMD"
date: "March 22, 2018"
output:
  html_document: default
  pdf_document: default
highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EnvironmentalSettingToolkit)
library(ggplot2)
```

```{r include=FALSE, echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
dataPlot <- function(plotDF, plotY, yLabel) {
  # Iterate observations, graphing a scatter plot for each station
  for (s in unique(plotDF$uid)) {
  ggplot(data = plotDF, aes(x = plotDF$date[which(plotDF$uid == s)], y = plotY[which(plotDF$uid == s)])) + geom_line() + geom_point() + 
      labs(x = "Year",
      y = yLabel,
      title = plotDF$name[which(plotDF$uid == s)][1])
  }
  # plot.new()
  # par(pin=c(6,6))
  # par(mfrow = c(length(unique(plotDF$uid)), 1)) # All station graphs in one page column
  # par(mar = c(4.5, 4.5, 1.5, 1.5)) # Graph margins in lines
  # for (s in unique(plotDF$uid)) {
  #   plot(
  #     CST1Data$date[which(plotDF$uid == s)],
  #     plotY[which(plotDF$uid == s)],
  #     pch = 16,
  #     col = "blue",
  #     ylab = yLabel,
  #     xlab = "date",
  #     cex.main=0.8,
  #     main = plotDF$name[which(plotDF$uid == s)][1]
  #   )
  # }
}
```
The draft Environmental Setting protocol calculates four types of climate metrics using the EnvironmentalSetting toolkit <https://github.com/nationalparkservice/EnvironmentalSetting_Toolkit>. 

* station-based temperature and precipitation
* station-based departures from normal temperature and precipitation
* grid-based departures from normal temperature and precipitation
* index-based precipitation and drought

The first two metric types are calculated for all stations that intersect the 30km areas of analysis for parks while the latter two metric types are calculated by area of analysis (AOA) bounding box extent. This notebook shows the first two metric types for stations around Agate Fossil Beds NM (AGFO). The climate stations and observation data for areas around and within National Parks are obtained from webservices maintained by NOAA's Regional Climate Centers (RCCs) from a system called ACIS - Applied Climate Information System <http://www.rcc-acis.org/index.html>.


```{r results = "asis", echo=FALSE}
metrics <- read.csv("inst/ESParameters.csv")[ ,1:2]

knitr::kable(metrics, type = "html")
```

# Station-Based Temperature and Precipitation Metrics
Examples shown are for stations within 20km of Agate Fossil Beds National Monument (AGFO). These are raw data that include missing values. 

## Temperature
``` {r message=FALSE, warning=FALSE, error=FALSE}
# Get stations collecting maxt, mint, pcpn, and gdd with data from last 20 years
agfoStations20km <- findStation(unitCode = "AGFO", distance = 20, climateParameters = list("maxt","mint","gdd"))
agfoStationsCurrent <- agfoStations20km [agfoStations20km$minDate >= "1998-02-10" & agfoStations20km$maxDate >= "2018-02-10" ,]
str(agfoStationsCurrent)

```

Data requests are made with the getWxObservations() function. Exploratory plots are shown in this notebook but are not included in the Toolkit functions.

```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST1: Hot Days (annual count)
CST1Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("maxt"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_90", metric = "CST1")
CST1Data$maxt_F_cnt_ge_90[order(CST1Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST1Data, aes(x = CST1Data$date, y = CST1Data$maxt_F_cnt_ge_90[order(CST1Data$name)], group=sort(CST1Data$name), color=sort(CST1Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST1Data$name)) +
      labs(x = "Year",
      y = "max temp >= 90 (annual count)",
      color = "Station",
      title = "CST 1: Days with maxt >= 90 (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r dpi=96, message=FALSE, warning=FALSE, error=FALSE}
# dataPlot(CST1Data,CST1Data$maxt_F_cnt_ge_90,"max temp >= 90 (annual count)")

```


```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST2: Cold Days (annual count)
CST2Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("maxt"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_le_32", metric = "CST2")
CST2Data$maxt_F_cnt_le_32[order(CST2Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST2Data, aes(x = CST2Data$date, y = CST2Data$maxt_F_cnt_le_32[order(CST2Data$name)], group=sort(CST2Data$name), color=sort(CST2Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST2Data$name)) +
      labs(x = "Year",
      y = "max temp <= 32 (annual count)",
      color = "Station",
      title = "CST 2: Days with maxt <= 32 (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
#dataPlot(CST2Data,CST2Data$maxt_F_cnt_le_32,"annual count (CST2)")

```

```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST3: Sub-freezing Days
CST3Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("mint"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_le_32", metric = "CST3")
CST3Data$mint_F_cnt_le_32[order(CST3Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST3Data, aes(x = CST3Data$date, y = CST3Data$mint_F_cnt_le_32[order(CST3Data$name)], group=sort(CST3Data$name), color=sort(CST3Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST3Data$name)) +
      labs(x = "Year",
      y = "min temp <= 32 (annual count)",
      color = "Station",
      title = "CST 3: Days with mint <= 32 (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
#dataPlot(CST3Data,CST3Data$mint_F_cnt_le_32,"annual count (CST3)")

```

```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST4: Days at or below 0 (annual count)
CST4Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("mint"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_le_0", metric = "CST4")
CST4Data$mint_F_cnt_le_0[order(CST4Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST4Data, aes(x = CST4Data$date, y = CST4Data$mint_F_cnt_le_0[order(CST4Data$name)], group=sort(CST4Data$name), color=sort(CST4Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST4Data$name)) +
      labs(x = "Year",
      y = "min temp <= 0 (annual count)",
      color = "Station",
      title = "CST 4: Days with mint <= 0 (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST5: Growing degree days (base temperature >= 32) (annual count)
CST5Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("gdd32"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_gt_0", metric = "CST5")
CST5Data$gdd32__cnt_gt_0[order(CST5Data$name)]

```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST5Data, aes(x = CST5Data$date, y = CST5Data$gdd32__cnt_gt_0[order(CST5Data$name)], group=sort(CST5Data$name), color=sort(CST5Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST5Data$name)) +
      labs(x = "Year",
      y = "growing degree days (annual count)",
      color = "Station",
      title = "CST 5: Growing degree days (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST6: Heating degree days (default base temperature >= 65) (annual count)
CST6Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("hdd"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_gt_0", metric = "CST6")
CST6Data$hdd__cnt_gt_0[order(CST6Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST6Data, aes(x = CST6Data$date, y = CST6Data$hdd__cnt_gt_0[order(CST6Data$name)], group=sort(CST6Data$name), color=sort(CST6Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST6Data$name)) +
      labs(x = "Year",
      y = "heating degree days (annual count)",
      color = "Station",
      title = "CST 6: Heating degree days (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST7: Cooling degree days (default base temperature >= 65) (annual count)
CST7Data <- getWxObservations(climateStations = unique(agfoStationsCurrent$uid), climateParameters = list("cdd"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_gt_0", metric = "CST7")
CST7Data$cdd__cnt_gt_0[order(CST7Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CST7Data, aes(x = CST6Data$date, y = CST7Data$cdd__cnt_gt_0[order(CST7Data$name)], group=sort(CST7Data$name), color=sort(CST7Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CST7Data$name)) +
      labs(x = "Year",
      y = "cooling degree days (annual count)",
      color = "Station",
      title = "CST 7: Cooling degree days (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

For the departure-based metrics, a new set of stations collecting average temperature is needed. 
```{r message=FALSE, warning=FALSE, error=FALSE}
agfoStationsBBoxAvgT <- findStation(unitCode = "AGFO", customBBox = "-104.19609745831781, 42.112886184389545, -103.2970796760243, 42.7314612519983", climateParameters = list("avgt"))
agfoStationsCurrentAvgT <- agfoStationsBBoxAvgT [agfoStationsBBoxAvgT$minDate >= "1998-02-10" & agfoStationsBBoxAvgT$maxDate >= "2018-02-10" ,]
str(agfoStationsCurrentAvgT)
```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST8 and 9: Annual count of days above/below 30-year climate normal (1981-2010)
CST8and9Source <- getWxObservations(climateStations = unique(agfoStationsCurrentAvgT$uid), climateParameters = list("avgt"), sdate="1998-02-10", edate="2018-02-10", duration="dly", interval = "dly", normal = "departure", metric = "CST8and9")
#CST8and9Data$avgt__F_departure
#str(CST8and9Source)
CST8and9Data <- getDepartureCounts(rawDepartures = CST8and9Source, duration = "yly", metric = "CST8and9")
str(CST8and9Data)
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
vals <- reshape2::melt(CST8and9Data, id.vars=c("date","name"), measure.vars = c("cntAboveNormal","cntBelowNormal"))
ggplot(vals, aes(x=date, y=as.numeric(value), color=variable, group=name)) + geom_point() + facet_wrap(~name) + labs(
  x = "Year",
  y = "departure (annual count)",
  color = "metric",
  title = "CST 8 and 9: Temperature departure from 30-year normal (1981-2010)"
  ) +
  theme(axis.text.x = element_text(
  angle = 90,
  vjust = 0.5,
  hjust = 1
  ))
# ggplot(data = CST8and9Data, aes(x = CST8and9Data$date)) + 
#   geom_line(aes(
#   y = CST8and9Data$cntAboveNormal,
#   group = CST8and9Data$name,
#   color = "above"
#   )) + geom_point(aes(
#   y = CST8and9Data$cntAboveNormal,
#   group = CST8and9Data$name,
#   color = "above"
#   )) + geom_line(aes(
#   y = CST8and9Data$cntBelowNormal,
#   group = CST8and9Data$name,
#   color = "below"
#   ))  +  geom_point(aes(
#   y = CST8and9Data$cntBelowNormal,
#   group = CST8and9Data$name,
#   color = "below"
#   )) + facet_wrap(~ CST8and9Data$name) +
#   labs(
#   x = "Year",
#   y = "departure (annual count)",
#   color = "Station",
#   title = "CST 8 and 9: Departure from 30-year normal (1981-2010)"
#   ) +
#   theme(axis.text.x = element_text(
#   angle = 90,
#   vjust = 0.5,
#   hjust = 1
#   ))

```

## Precipitation
For the departure-based metrics, a new set of stations collecting precipitation is needed. Notice the additional station discovered and that the three stations do not collect snow depth.
```{r message=FALSE, warning=FALSE, error=FALSE}
agfoStationsBBoxPcpn <- findStation(unitCode = "AGFO", customBBox = "-104.19609745831781, 42.112886184389545, -103.2970796760243, 42.7314612519983", climateParameters = list("pcpn","snow"))
agfoStationsCurrentPcpn <- agfoStationsBBoxPcpn [agfoStationsBBoxPcpn$minDate >= "1998-02-10" & agfoStationsBBoxPcpn$maxDate >= "2018-02-10" ,]
str(agfoStationsCurrentPcpn)
```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP1: Heavy precipitation days (annual count)
CSP1Data <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("pcpn"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_1.0", metric = "CSP1")
CSP1Data$pcpn_in_cnt_ge_1.0[order(CSP1Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
ggplot(data = CSP1Data, aes(x = CSP1Data$date, y = CSP1Data$pcpn_in_cnt_ge_1.0[order(CSP1Data$name)],  group=sort(CSP1Data$name), color=sort(CSP1Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CSP1Data$name)) +
      labs(x = "Year",
      y = "pcpn >= 1.0 inch (annual count)",
      color = "Station",
      title = "CSP 1: Days with pcpn >= 1.0 inch (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP2: Extreme precipitation days (annual count)
CSP2Data <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("pcpn"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_2.0", metric = "CSP2")
CSP2Data$pcpn_in_cnt_ge_2.0[order(CSP2Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
ggplot(data = CSP2Data, aes(x = CSP2Data$date, y = CSP2Data$pcpn_in_cnt_ge_2.0[order(CSP2Data$name)], group=sort(CSP2Data$name), color=sort(CSP2Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CSP2Data$name)) +
      labs(x = "Year",
      y = "pcpn >= 2.0 inch (annual count)",
      color = "Station",
      title = "CSP 2: Days with pcpn >= 2.0 inches (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP3: Micro-drought - runs of 7 or more days with less than 0.01in precip (annual count)
CSP3Source <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("pcpn"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "run_le_0.01", metric = "CSP3")
CSP3Source$pcpn_in_runYear[order(CSP3Source$name)]
CSP3Data <- getRunCounts(rawCounts = CSP3Source, runLength = 7, metric = "CSP3")
CSP3Data$cntGERunLength_7[order(CSP3Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
# ggplot(data = CSP3Data, aes(x = CSP3Data$date, y = CSP3Data$cntGERunLength_7, group=CSP3Data$name, color=CSP3Data$name)) + geom_line() + geom_point() + facet_wrap(~ CSP3Data$name) +
#       labs(x = "Year",
#       y = "run >= 7 days (annual count)",
#       color = "Station",
#       title = "CSP 3: Micro-drought - runs of 7 or more days with less than 0.01in precipitation") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

vals <- reshape2::melt(CSP3Data, id.vars=c("date","name"), measure.vars = c("cntGERunLength_7"))
ggplot(vals, aes(x=date, y=as.numeric(value), color=name, group=name)) + geom_point() + facet_wrap(~name) + labs(
  x = "Year",
  y = "run >= 7 days (annual count)",
  color = "metric",
  title = "CSP 3: Micro-drought (runs of 7 or more days with less than 0.01in precipitation)"
  ) +
  theme(axis.text.x = element_text(
  angle = 90,
  vjust = 0.5,
  hjust = 1
  ))
```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP4: Measurable snow days (annual count)
CSP4Data <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("snow"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_0.1", metric = "CSP4")
CSP4Data$snow_in_cnt_ge_0.1[order(CSP4Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CSP4Data, aes(x = CSP4Data$date, y = CSP4Data$snow_in_cnt_ge_0.1[order(CSP4Data$name)], group=sort(CSP4Data$name), color=sort(CSP4Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CSP4Data$name)) +
      labs(x = "Year",
      y = "snow >= 0.1 inch (annual count)",
      color = "Station",
      title = "CSP 4: Days with measurable snow >= 0.1 inches (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP5: Moderate snow days (annual count)
CSP5Data <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("snow"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_3.0", metric = "CSP5")
CSP5Data$snow_in_cnt_ge_3.0[order(CSP5Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CSP5Data, aes(x = CSP5Data$date, y = CSP5Data$snow_in_cnt_ge_3.0[order(CSP5Data$name)], group=sort(CSP5Data$name), color=sort(CSP5Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CSP5Data$name)) +
      labs(x = "Year",
      y = "snow >= 3.0 inches (annual count)",
      color = "Station",
      title = "CSP 5: Days with moderate snow >= 3.0 inches (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CSP6: Heavy snow days (annual count)
CSP6Data <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("snow"), sdate="1998-02-10", duration="yly", interval = "yly", reduceCodes = "cnt_ge_5.0", metric = "CSP6")
CSP6Data$snow_in_cnt_ge_5.0[order(CSP6Data$name)]
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
  ggplot(data = CSP6Data, aes(x = CSP6Data$date, y = CSP6Data$snow_in_cnt_ge_5.0[order(CSP6Data$name)], group=sort(CSP6Data$name), color=sort(CSP6Data$name))) + geom_line() + geom_point() + facet_wrap(~ sort(CSP6Data$name)) +
      labs(x = "Year",
      y = "snow >= 5.0 inches (annual count)",
      color = "Station",
      title = "CSP 6: Days with heavy snow >= 5.0 inches (annual count)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r message=FALSE, warning=FALSE, error=FALSE}
# Get CST8 and 9: Annual count of days above/below 30-year climate normal (1981-2010)
CSP7and8Source <- getWxObservations(climateStations = unique(agfoStationsCurrentPcpn$uid), climateParameters = list("pcpn"), sdate="1998-02-10", edate="2018-02-10", duration="dly", interval = "dly", normal = "departure", metric = "CSP7and8")
#CST8and9Data$avgt__F_departure
#str(CST8and9Source)
CSP7and8Data <- getDepartureCounts(rawDepartures = CSP7and8Source, duration = "yly", metric = "CSP7and8")
str(CSP7and8Data)
```
```{r echo=FALSE, dpi=96, message=FALSE, warning=FALSE, error=FALSE}
vals <- reshape2::melt(CSP7and8Data, id.vars=c("date","name"), measure.vars = c("cntAboveNormal","cntBelowNormal"))
ggplot(vals, aes(x=date, y=as.numeric(value), color=variable, group=name)) + geom_point() + facet_wrap(~name) + labs(
  x = "Year",
  y = "departure (annual count)",
  color = "metric",
  title = "CSP 7 and 8: Precipitation departure from 30-year normal (1981-2010)"
  ) +
  theme(axis.text.x = element_text(
  angle = 90,
  vjust = 0.5,
  hjust = 1
  ))
```